---
title: "Project_Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Project_Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Project Prompt

Use the SMS survey and co-located sensor rainfall data to identify mid-season dry-spells of potential agricultural significance. For example, two or more consecutive rainless weeks close to maize flowering or grain-filling times. Can you detect a signficant drop in farmers’ anticipated harvest amounts around such events (we might expect a sudden drop in expected harvest amounts)? How about a significant change in the rate of reduction in farmers’ reported maize storage levels fall (if farmers are concerned about crop loss, they might slow their rate of consumption of stored food)?


# Description
```{r setup}
library(SouthAfricaDrought)
library(geospaar)
library(sf)
library(rgdal)
library(raster)

#setwd("C:/Users/Emily/Documents/Fall_2020/Geospatial_R/SouthAfricaDrought")
#Load farmer data
#fs <- system.file("data-raw/SouthAfrica_Boundary.shp", package = "SouthAfricaDrought") %>% st_read()
#boundary <- roads <- st_read(dsn = fs)

```

# Methods

- Deciles - recommended by drought.gov

- NOAA Drought Index - also recommended by drought.gov

- Temperature above monthly average

- EVI - Remote Sensing indicator for vegetation health


# Data Visualization

# Results
