---
title: "Project_Overview"
output: html_document
runtime: shiny
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Project Overview
## Authors: Emily Evenden and Maseeng Masitha
## Date: November 23rd, 2020

## Summary

We aim to identify bouts of drought in South Africa over the summer season of November, 2019, to March, 2020. Drought will be determined by persisting stretches of no rainfall. We will use two drought indices which rely solely on rainfall data: the Deciles Index and the NOAA Drought Index to identify areas of South Africa expericning drought. Through this project, we hope to program easy drought metrics for South Africa which could be used to determine early warnings of low crop yields.

# Primary Objectives

- Make long-term and short-term rainfall data available to the user in both monthly and weekly time intervals
- Program two drought indices
- Identify areas in South Africa that experienced drought last summer (2019-2020).

# Approach and Method
## Data

This data contains five datasets:

  1. SouthAfrica_Boundary - a vector file containing the boundary of South Africa provided by the Humanitarian Data Exchange (available here: https://data.humdata.org/dataset/south-africa-admin-level-1-boundaries)
  
  2. SA_MODIS_Monthly_Mean_Temperature - a raster stack of the average monthly temperature (in Celsius) for South Africa. This dataset is not currently used in any of the indices we created, but is available for future use if more indicies are added.
  
  3. & 4. SA_Chirps_Weekly_Mean_Precip & SA_Chirps_Monthly_Mean_Precip - these datasets are two different aggregations of the daily Chirps precipitation data. Obviously, the weekly data is aggregated by week while the monthly data is aggregated by month. The unit of this dataset is mm of rainfall per day.
  
  5. SA_Chirps_10yr_Monthly_Mean_Precip - this datasets is also dervied from the Chirps daily precipitation dataset. Each raster layer in this dataset shows the ten year average of rainfall for that month. The data range for this data is April 1, 1987 to March 31, 2018. 

We obtained datasets for temperature and precipitation from Google Earth Engine (GEE) database. Specifically, we utilized the "Chirps Daily: Climate Hazards Group Infrared Precipitation" dataset (available here: https://developers.google.com/earth-engine/datasets/catalog/UCSB-CHG_CHIRPS_DAILY) and the "Terra Land Surface Temperature and Emissivity Daily Global 1km" dataset (available here: https://developers.google.com/earth-engine/datasets/catalog/MODIS_006_MOD11A1). These datasets were filtered by data and location and reduced using the mean function via the Javascript Code Editor available online through GEE. Once selected, the data was exported to Google Drive.  

Finally data preprocessing was conducted using R. Following this, the datasets were clipped in R to the extent of the study area, South Africa. A function for calculating mean monthly amounts for the variables was applied on temperature and precipitation datasets and units converted where appropriate: Kelvin to degree Celsius. The next geoprocessing of the datasets was done in raster layers, and included calculating the mean of pixels throughout the months to produce a new output file.

```{r setup, error=TRUE, echo = FALSE, message = FALSE}

library(SouthAfricaDrought)
library(geospaar)
library(sf)
library(rgdal)
library(raster)
library(shiny)
library(shinydashboard)
library(leaflet)
library(ggplot2)

afalb <- paste0("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 ", "+y_0=0 +ellps=WGS84 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs")

#Load South Africa boundary and reprojects to Hartebeesthoek94
boundary <- system.file("extdata/SouthAfrica_Boundary.shp", package = "SouthAfricaDrought") %>% st_read() %>% st_transform(x = ., crs = afalb)

#Load long-term monthly data and reprojects custom projection for SA
lt_monthly_prec <- system.file("extdata/SA_30yr_LT_Mean_Precip.tif", package = "SouthAfricaDrought") %>% stack() %>% projectRaster(., crs = afalb)

#Rename according to month
names(lt_monthly_prec) <- c('Nov_87_17', 'Dec_87_17', 'Jan_88_18', 'Feb_88_18', 'March_88_18')
                              
#Load summer monthly data and reprojects custom projection for SA
monthly_prec <- system.file("extdata/SA_Chirps_Monthly_Mean_Precip.tif", package = "SouthAfricaDrought") %>% stack() %>% projectRaster(., crs = afalb)

#Rename layers accoridng to their month
                                                                                                                          names(monthly_prec) <- c('Nov_2019', 'Dec_2019', 'Jan_2020', 'Feb_2020', 'March_2020')
plot(monthly_prec)                                                                                                                              
#Load summer weekly precipitation values (mm) and reprojects custom projection for SA
weekly_prec <- system.file("extdata/SA_Chirps_Weekly_Mean_Precip.tif", package = "SouthAfricaDrought") %>% stack() %>% projectRaster(., crs = afalb)

#Rename precipitation layers according to the data range of each image
week_names <- c('week1_2019-10-27_2019-11-02',
                        'week2_2019-11-17_2019-11-23',
                        'week3_2019-11-24_2019-11-30',
                        'week4_2019-12-01_2019-12-07',
                        'week5_2019-12-08_2019-12-14',
                        'week6_2019-12-15_2019-12-21',
                        'week7_2019-12-22_2019-12-28',
                        'week8_2019-12-29_2019-01-05',
                        'week9_2019-01-06_2019-01-12',
                        'week10_2019-01-13_2019-01-19',
                        'week11_2019-01-20_2019-01-26',
                        'week12_2019-01-27_2019-02-02',
                        'week13_2019-02-03_2019-02-09',
                        'week14_2019-02-10_2019-02-16',
                        'week15_2019-02-17_2019-02-23',
                        'week16_2019-02-24_2019-03-02',
                        'week17_2019-03-10_2019-03-16',
                        'week18_2019-03-17_2019-02-23',
                        'week19_2019-03-24_2019-03-30',
                        'week20_2019-03-31_2019-04-06',
                        'week21_2019-04-07_2019-04-13',
                        'week22_2019-04-14_2019-04-20',
                        'week23_2019-02-21_2019-02-27')

names(weekly_prec) <- week_names
weekly_prec
```

## Methods


# 1. Deciles Index

The Deciles index compares precipitation from a specific time interval with precipitation from a historic time interval (usually 30 years or more). This index only uses precipitation data. The decile method divides each month of historic precipitation data  into ten equal parts. These categories can then be used to assess the new time interval. The decile ranges can be categorized as such:

  Deciles 1-2: Much below Normal
  Deciles 3-4: Below Normal
  Deciles 5-6: Near Normal
  Deciles 7-8: Above Normal
  Deciles 8-9: Much above Normal

Information source: https://agrimetsoft.com/faq/What%20is%20DI(Deciles%20Index)

```{r, error= TRUE, echo = FALSE, message=FALSE}
#write out code here and explain
#Deciles
q1 <- quantile(monthly_prec, c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), na.rm=TRUE)

q1

#NOAA Drough Index
```



# 2. NOAA Drought Index

The NOAA Drought Index is also a precipitation based index. This index utilized weekly precipitation data for the time interval of interest. In this case, the mean precipitation for an eight-week interval is calculated and then compared with the subsequent ninth week. If the ninth week is greater than 60% of the precipitation, it's classified as having little to no water stress. An area is considered no longer stressed once the rainfall returns to 60% or more. 

Information source: https://www.droughtmanagement.info/noaa-drought-index-ndi/

```{r, error= TRUE, echo = FALSE, message=FALSE}

result_list <- list()

noaa_stack <- SouthAfricaDrought::noaa_drought(weekly_prec, result_list) %>% stack()

noaa_class_stack <- SouthAfricaDrought::noaa_class(noaa_stack)

names(noaa_class_stack) <- c('Week 9',
                             'Week 10',
                             'Week 11',
                             'Week 12',
                             'Week 13',
                             'Week 14',
                             'Week 15',
                             'Week 16',
                             'Week 17',
                             'Week 18',
                             'Week 19',
                             'Week 20',
                             'Week 21',
                             'Week 22',
                             'week 23')
plot(noaa_class_stack)
```



# Data Visualization

```{r, echo = FALSE, message=FALSE}
#Display final maps of drought -- times series of drought areas?
pal <- colorFactor(c("red" ), c(1), na.color = "transparent")

shinyApp(ui <- fluidPage(
  
  titlePanel("Weekly Raster"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      sliderInput(inputId = "slider", label = "Week", min = 0, max = 14, value = 1)
    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output:
      leafletOutput("Raster", width = "100%")
    )
  )),


server <- function(input, output, session) {

  output$Raster <- renderLeaflet({
    
    leaflet() %>% 
      
      addTiles() %>% 
      
      setView(lat= -29, lng=24, zoom = 5) %>% 
      
      addRasterImage(df_filtered(), color = pal) %>% 
      
      addLegend("bottomright", pal = pal, values = c(0, 1), labels = c("", "Area of Water Stress"), opacity = 1)

  })
  
  df_filtered <- reactive({
    
    !is.na(noaa_class_stack[[input$slider]])
    
    })
  
  observe({
    leafletProxy(mapId = "Raster") %>% clearImages() %>% addRasterImage(df_filtered(), color = pal, opacity = 0.6)
  })
  
  options = list(height = 800)

})

```

# Results

# Contribution
## Maseeng

## Emily